Fast gravitational wave parameter estimation without compromises

### 2023-07-21 src/tex/ms.tex
Add additional run statistics

This is precisely why \textsc{flowMC} is necessary for our approach. Since every
accepted sample generated using the normalizing flow as proposal is independent,
and the sampling from the normalizing flow is an embarrassingly parallel
process, the name of the game then becomes training the normalizing flow as fast as
possible. Having many chains helps to produce the necessary amount of samples
for training the normalizing flow, which means now we can meaningfully harvest
the benefit of having large throughput of GPU. Without either of this two,
our approach would be significantly less performant.}

@@

To demonstrate the performance of our parameter estimation pipeline, we apply it
to two real LIGO-Virgo events: GW150914 and GW170817. We use the priors shown in
Table \ref{tab:parameters}, and take 4 s of data sampled at 2048 Hz starting at
20 Hz for GW150914, and 128 s of data sampled at 4096 Hz starting at 23 Hz for
GW170817; strain data and PSDs for both events are fetched from GWOSC
\citep{GWOSC}. We use the same normalizing flow model as the injection-recovery
study. For our specific choice of sampler settings, we produce ${\sim}$2500 and
3500 \emph{effective samples}\footnote{Effective samples here refers to the
number of independent samples, which is the total number of generated samples
divided by their correlation length; we compute the effective sample size using
\textsc{arviz} \citep{arviz_2019},
\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}. } for
GW150914 and GW170817 respectively. \corr{Running on a node with an
\textsc{Nvidia} 80GB A100 GPU and 16 intel skylake CPUs} \footnote{Note that the
CPU time is hard to benchmark in \textsc{Jax} due to JIT compilation and
asynchronous dispatch. Overall, we find 4 CPUs to be sufficient for our use
case. We chose 16 because of the ease of use of the local cluster.}, the wall
time for both events is around 3 minutes. Most of this time is spent on
just-in-time (JIT) compilation of the code; the actual sampling time is only
${\sim}40$ s. We pre-compute the reference waveform parameters used to
hetetrodyne the likelihood for the two events, which is omitted in the wall-time
calculation.

@@

For comparison, we produce equivalent runs with \textsc{Bilby}, using the same
exact data and priors.  We use the \textsc{dynesty} sampler
\citep{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
in the configurations files available on github \url{https://github.com/kazewong/TurboPE/tree/main/src}.  We
carry out these runs using \textsc{parallel Bilby} (\textsc{pBilby})
\citep{Smith:2019ucc} to distribute the computation over 400 Intel Skylake CPUs
for each event.  For the specific settings chosen, the wall-time duration of
each run was ${\sim}2$ h for GW150914 and ${\sim}1$ day for GW170817.

% Describe the distance between flowMC result and Bilby result using
% Jensen-Shannon divergence, scipy has a function to compute that in
% scipy.spatial.distance

>>>>

This is precisely why \textsc{flowMC} is necessary for our approach. Since every
accepted sample generated using the normalizing flow as proposal is nearly independent,
and the sampling from the normalizing flow is an embarrassingly parallel
process, the name of the game then becomes training the normalizing flow as fast as
possible. Having many chains helps to produce the necessary amount of samples
for training the normalizing flow, which means now we can meaningfully harvest
the benefit of having large throughput of GPU. Without either of this two,
our approach would be significantly less performant.}

@@

To demonstrate the performance of our parameter estimation pipeline, we apply it
to two real LIGO-Virgo events: GW150914 and GW170817. We use the priors shown in
Table \ref{tab:parameters}, and take 4 s of data sampled at 2048 Hz starting at
20 Hz for GW150914, and 128 s of data sampled at 4096 Hz starting at 23 Hz for
GW170817; strain data and PSDs for both events are fetched from GWOSC
\citep{GWOSC}. We use the same normalizing flow model as the injection-recovery
study. For our specific choice of sampler settings, we produce ${\sim}$2500 and
3500 \emph{effective samples}\footnote{Effective samples here refers to the
number of independent samples, which is the total number of generated samples
divided by their correlation length; we compute the effective sample size using
\textsc{arviz} \citep{arviz_2019},
\url{https://python.arviz.org/en/stable/api/generated/arviz.ess.html}. } for
GW150914 and GW170817 respectively. \corr{Running on a node with an
\textsc{Nvidia} 80GB A100 GPU and 16 intel skylake CPUs} \footnote{Note that the
CPU time is hard to benchmark in \textsc{Jax} due to JIT compilation and
asynchronous dispatch. Overall, we find 4 CPUs to be sufficient for our use
case. We chose 16 because of the ease of use of the local cluster.}, the wall
time for both events is around 3 minutes. Most of this time is spent on
just-in-time (JIT) compilation of the code; the actual sampling time is only
${\sim}40$ s. We pre-compute the reference waveform parameters used to
hetetrodyne the likelihood for the two events, which is omitted in the wall-time
calculation. \corr{In terms of specific configuration of the sampler, we used
1000 chains for both of the events. We run both the tuning phase and the
production phase for 20 loops each. Within each phase, the global sampler and
local sampler are stepped 200 times. This translates into a total number of
likelihood evaluation is $1.6*10^7$. \footnote{While in principle adding more
chains makes the algorithm less efficient in terms of acceptance/evaluation
ratio, since GPUs can evaluate a fixed number of instruction per cycle, we can
keep adding chains without paying extra computational cost before we saturate
the GPU. This results in an improvement in wall time while being less efficient.
We decide to choose a setting that improve wall time instead of efficiency.} The
acceptance rate of the global proposal is about $2\%$ and $5\%$ for GW150914 and
GW170817 respectively. Note that the setting is chosen to run longer than
necessary to ensure convergence, we are currently pursing tuning studies to
understand trade-offs in the algorithms and improve the efficiency further.}

@@

For comparison, we produce equivalent runs with \textsc{Bilby}, using the same
exact data and priors.  We use the \textsc{dynesty} sampler
\citep{2020MNRAS.493.3132S,dynesty}, with 1000 live points and other settings as
in the configurations files available on github \url{https://github.com/kazewong/TurboPE/tree/main/src}.  We
carry out these runs using \textsc{parallel Bilby} (\textsc{pBilby})
\citep{Smith:2019ucc} to distribute the computation over 400 Intel Skylake CPUs
for each event. For the specific settings chosen, the wall-time duration of
each run was ${\sim}2$ h for GW150914 and ${\sim}1$ day for GW170817.
\corr{We have also benchmarked the wall-time performance }
